Index measures for oak decline severity using phenotypic descriptors

### 2020-02-07 manuscript/manuscript.Rmd
alterations from Johns comments

---
output: 
  word_document: default
  pdf_document: default
header-includes:
  - \linespread{2}
  - \usepackage{lineno}
  - \linenumbers
bibliography: references.bib
---

@@

Assigning discrete decline status labels neglects the true spectrum of decline severity encountered in the field.
There is a growing need for sensitive, continuous measures of tree decline that accurately reflect the decline spectrum [@pontius2014].
This has the potential to greatly improve monitoring, enable the early detection of decline for accurate forecasting and inform forestry management practices.
Consistent long term monitoring has been highlighted as a key requirement for understanding the threats to oak and safeguarding its future [@hartmann2018].
Coupled with this, monitoring methods should take advantage of modern data analysis techniques whilst still retaining practical simplicity in the field.

@@

All cardinal assessments were averaged to give a single value per tree.
The crown density ($CD$) values erroneously collected at Attingham and Langdale were converted to crown transparency ($t = 100 - CD$)
All descriptors collected in cm were converted to mm except for DBH which was converted to m.
The tap tests that were positive for hollow stem tissue were converted to present and absent then renamed dead stem tissue.
The canopy closure descriptor was converted from yes and no to present and absent.
Finally, the stem symptoms sizes for active bleeds, black staining and callused wounds were averaged and a frequency per tree collected.
After preparation there were a total of 31 raw descriptors for analysis.

@@

To investigate the descriptor contributions to the decline indexes, supervised random forest regression models were trained for each decline index using the phenotypic descriptors for all 174 trees.
These models will subsequently be referred to as $M_{PDI}$ and $M_{DAI}$ for the PDI and DAI respectively. 
The optimal number of trees ($ntree$) and number randomly of variables ($mtry$) parameters for random forest model were tuned prior to training using the values of 10, 100, 1,000, 10,000 for ntree and a range of 2-26 at intervals of 4 for the mtry parameter. 
The combination parameters giving the lowest percent mean absolute error (PMAE) were selected and set as follows: for the PDI ($M_{PDI}$): `r glue('$ntree$ = {prettyNum(PDI_rf_tune_params$ntree,big.mark = ',')}, $mtry$ = {PDI_rf_tune_params$mtry}, PMAE = {PDI_rf_tune_params$PMAE %>% signif(digits = 3)}%')`; for the DAI ($M_{DAI}$): `r glue('$ntree$ = {prettyNum(DAI_rf_tune_params$ntree,big.mark = ',')}, $mtry$ = {DAI_rf_tune_params$mtry}, PMAE = {DAI_rf_tune_params$PMAE %>% signif(digits = 3)}%')` (Supplementary figure 2).
Overall descriptor importance contributions to each of the indexes was measured in each model using the mean decrease in accuracy measure.

>>>>

---
output: 
  word_document: default
  pdf_document: default
fontsize: 12pt
header-includes:
  - \linespread{2}
  - \usepackage{lineno}
  - \linenumbers
bibliography: references.bib
---

@@

Assigning discrete decline status labels neglects the true spectrum of decline severity encountered in the field.
There is a growing need for sensitive, continuous measures of tree decline that accurately reflect the decline severity spectrum [@pontius2014].
This has the potential to greatly improve monitoring, enable the early detection of decline for accurate forecasting and inform forestry management practices.
Consistent long term monitoring has been highlighted as a key requirement for understanding the threats to oak and safeguarding its future [@hartmann2018].
Coupled with this, monitoring methods should take advantage of modern data analysis techniques whilst still retaining practical simplicity in the field.

@@

All cardinal assessments were averaged to give a single value per tree.
The crown density ($CD$) values collected at Attingham and Langdale were converted to crown transparency ($t = 100 - CD$)
All descriptors collected in cm were converted to mm except for DBH which was converted to m.
The tap tests that were positive for hollow stem tissue were converted to present and absent then renamed dead stem tissue.
The canopy closure descriptor was converted from yes and no to present and absent.
Finally, the stem symptoms sizes for active bleeds, black staining and callused wounds were averaged and a frequency per tree totalled.
After preparation there were a total of 31 raw descriptors for analysis.

@@

To investigate the descriptor contributions to the decline indexes, supervised random forest regression models were trained for each decline index using the phenotypic descriptors for all 174 trees.
These models will subsequently be referred to as $M_{PDI}$ and $M_{DAI}$ for the PDI and DAI respectively. 
The optimal number of trees ($ntree$) and number randomly of variables ($mtry$) parameters for random forest model were tuned prior to training using the values of 10, 100, 1,000, 10,000 for ntree and a range of 2-26 at intervals of 4 for the mtry parameter. 
The combination parameters giving the lowest percent mean absolute error (PMAE) were selected and set as follows: for the PDI ($M_{PDI}$): `r glue('$ntree$ = {prettyNum(PDI_rf_tune_params$ntree,big.mark = ',')}, $mtry$ = {PDI_rf_tune_params$mtry}, PMAE = {PDI_rf_tune_params$PMAE %>% signif(digits = 3)}%')`; for the DAI ($M_{DAI}$): `r glue('$ntree$ = {prettyNum(DAI_rf_tune_params$ntree,big.mark = ',')}, $mtry$ = {DAI_rf_tune_params$mtry}, PMAE = {DAI_rf_tune_params$PMAE %>% signif(digits = 3)}%')` (Supplementary figure 2).
Overall descriptor importance contributions to each of the indexes was measured in each model using the mean decrease in accuracy measure with higher values considered to have a higher overall contribution to the decline index.
